### 2015-06-06 LOI/inc2.tex
My edits.

\underline{\textbf{Objectives:}}
Exascale scientific discovery will be severely bottlenecked without sufficient
new research into managing and storing the large amounts of data that will be
produced during the simulation, and analyzed for months afterwards.  Our goal
in this project is to address the associated I/O and storage challenges on the
future storage landscape, and expedite insights into mission critical
scientific processes. To that end, we will build on the capabilities offered
by the Sirocco peer-to-peer file system under development at Sandia and the
object storage and annotation expertise of UC Santa Cruz to develop a unique
application aware data management scheme. It will build on the existing
multi-tier storage aware approach offered by Sirroco to incorporate new
functionality and extended object metadata and APIs for 1) matching available
storage bandwidth across the storage hierarchy with data importance employing
application aware data compression (both lossy with specified error bounds and
lossless), 2) a selectable performance/quality tradeoff when reading data, and
3) incorporate reader prioritization for data annotation, placement, and data
quality when writing data.  The major challenge in this project is metadata
management connecting different object quality levels and connecting the user
with the best possible outcomes given the current system state. The new user
APIs required to interact with this rich metadata system will drive effective
use of the entire middleware and storage stack. The project is compriesed of a
team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL,
UCSC) and storage (UCSC), and connect and coordinate these key storage
components in a seamless fashion.

Our approach can most easily be understood by considering a road navigation
program. Similar to the navigation program, we will offer a coarse grained,
fast overview of a data set with progressively more detail in areas of interest
noted by data features and less data in areas without. These progressively
more detailed data views require more storage space and time to retrieve. By
incorporating middleware informed of both how to identify relevant data
features and the intended use locations, it will be able to selectively store
data at different quality levels, with bounded errors in the case of lossy
compression, and allow a user to select what data to retrieve by specifying
both an acceptable retrieval timeframe and quality required. If the request
cannot be fulfilled, the storage system can negotiate with the user providing
information about the quality and timeframes available for a subsequent
request.

@@

\underline{\textbf{Key Technical Approach:}}
Our overall technical approach is based on an application-aware runtime
realization of tradeoffs in data representation, data placement and data
access. We will allow users to ``plug-in'' their knowledge about the data, not
as bytes but as motifs, allowing the I/O and storage system to understand user
intentions and the relationships between data objects, as well as data access
and transport patterns. This will facilitate efficient mapping of data from
user space, such as a multi-dimensional array, onto various storage tiers.
Rather than focus on simple data compression, we will incorporate a multitude
of techniques to reduce data on the faster yet smaller storage tiers, and keep
the less reduced information in the lower storage tiers.

We will support additional data access modes as well. \textbf{First}, within a
given accuracy bounds, offer a mode for re-computation from data stored in a
``fast'' tier to reduce data latencies. \textbf{Second}, exploratory analysis
operations frequently entail an overall data set view followed by targeted data
exploration based on identified features. We will offer support for a
configurable data access mode to support these sorts of analysis accesses. An
``overview'' access mode that gives a quick, approximate within error bounds,
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the complete, detailed data set from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. At the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.
\textbf{Third}, to ensure available storage for subsequent operations, we will
offer automatic data migration based on user annotations for required data
lifetimes using monitoring and learning techniques. Unlike existing approaches,
this will be tempered both by the user annotations and through learned access
patterns.  While past access patterns may not indicate future access because
the simulation run purpose may have changed, we are focused on scalability
where runs are subsequently larger as the simulation prepares for a capability
run.  By learning from the output and access patterns during this run sequence,
we can accurate decide how to place and organize data for the critical
capability runs. Fourth, we anticipate storing multiple data copies, each
compressed in different ways according to the underlying media, some of these
copies will disappear based on storage pressures, but data persistence will be
maintained according to user specifications. Assuming a relatively low latency
cache layer before a tape system, we can offer exploratory data access
reserving pulling data from tape to just the data required. This will save
scientists time and make data stored on tape usable without long delays.

>>>>

\underline{\textbf{Objectives:}}
Exascale scientific discovery will be severely bottlenecked without sufficient
new research into managing and storing the large amounts of data that will be
produced during the simulation, and analyzed for months afterwards.  Our goal
in this project is to address the associated I/O and storage challenges on the
future storage landscape, and expedite insights into mission critical
scientific processes. To that end, we will build on the capabilities offered
by ADIOS that provides I/O abstractions, 
the Sirocco peer-to-peer file system under development at Sandia and the
object storage and annotation expertise of UC Santa Cruz to develop a unique
application-aware data management scheme. It will center upon
a multi-tier storage-aware approach and incorporate new
functionalities and extended object metadata and APIs for 1) matching
storage capability across the storage hierarchy with data importance employing
application-aware data compression (both lossy with specified error bounds and
lossless), 2) a selectable performance/quality tradeoff when reading data, and
the ability to recompute data if needed, and
3) incorporate reader prioritization for data annotation, placement, and data
quality when writing data.  The major challenge in this project is metadata
management connecting different object quality levels and enabling the user
with the best possible outcomes given the user intent and the storage system state. The new user
APIs required to interact with this rich metadata system will drive effective
use of the entire middleware and storage stack. The project is compriesed of a
team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL,
UCSC) and storage (UCSC), and connect and coordinate these key storage
components in a seamless fashion.

Our approach can most easily be understood by considering a road navigation
program. Similar to the navigation program, we will offer a coarse grained,
fast overview of a data set with progressively more detail in areas of interest
such as those contain features, and less detail in areas without. These progressively
more detailed data views require more storage space and time to retrieve. By
incorporating middleware informed of either the criteria for identifying areas of interest
or the intended use locations, it will be able to selectively store
data at different quality levels, with bounded errors in the case of lossy
compression, and allow a user to select what data to retrieve by specifying
both an acceptable retrieval timeframe and error bound. 
%If the request
%cannot be fulfilled, the storage system can negotiate with the user providing
%information about the quality and timeframes available for a subsequent
%request.
%%%%%%%%%%%% I am not sure if we should do this type of expert system in the project.
The user may also specify guidlines that are funneled through I/O middle and 
advise how to react if the timeframe and error bound cannot be accepted.

@@

\underline{\textbf{Key Technical Approach:}}
Our overall technical approach is based on an application-aware runtime
realization of tradeoffs in data representation, data placement and data
access. We will allow users to be able to ``plug-in'' their knowledge 
about the data, not as bytes but as motifs, thus allowing the I/O and 
storage system to understand user
intentions and the relationships between data objects, as well as data access
and transport patterns. This will facilitate efficient mapping of data from
user space onto various storage tiers, and application-guided
data reductions.

The I/O and storage system produced from the project will
enable highly flexible data access modes that are common in data analytics. 
\textbf{First}, data re-computation is allowed to re-generate
data with higher accuracy, if the associated computing and I/O cost is lower
than retrieving the target data from a ``slow'' tier. 
\textbf{Second}, exploratory analysis
operations frequently entail an overall data set view followed by targeted data
exploration on extracted features. 
In particular, the ``overview'' access mode gives a quick but approximate
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the high-accuracy data from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. In the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.
\textbf{Third}, to ensure available storage for subsequent operations, we will
offer automatic data migration based on user annotations for required data
lifetimes using monitoring and learning techniques. Unlike existing approaches,
this will be tempered both by the user annotations and through learned access
patterns.  While past access patterns may not indicate future access because
the simulation run purpose may have changed, we are focused on scalability
where runs are subsequently larger as the simulation prepares for a capability
run.  By learning from the output and access patterns during this run sequence,
we can accurate decide how to place and organize data for the critical
capability runs. Fourth, we anticipate storing multiple data copies, each
compressed in different ways according to the underlying media, some of these
copies will disappear based on storage pressures, but data persistence will be
maintained according to user specifications. Assuming a relatively low latency
cache layer before a tape system, we can offer exploratory data access
reserving pulling data from tape to just the data required. This will save
scientists time and make data stored on tape usable without long delays.
